%\label{Bibliography}

\bibliographystyle{unsrtnat}
\addtotoc{BIBLIOGRAPHY}

{\scriptsize\setstretch{1.0}
\begin{thebibliography}{22}

\bibitem{engle1982} R.~F. Engle, ``Autoregressive conditional heteroscedasticity with estimates of the variance of United Kingdom inflation,'' \emph{Econometrica}, vol.~50, no.~4, pp.~987--1007, 1982.

\bibitem{bollerslev1986} T.~Bollerslev, ``Generalized autoregressive conditional heteroskedasticity,'' \emph{Journal of Econometrics}, vol.~31, pp.~307--327, 1986.

\bibitem{parkinson1980} M.~Parkinson, ``The extreme value method for estimating the variance of the rate of return,'' \emph{Journal of Business}, vol.~53, no.~1, pp.~61--65, 1980.

\bibitem{garman1980} M.~B. Garman and M.~J. Klass, ``On the estimation of security price volatilities from historical data,'' \emph{Journal of Business}, vol.~53, no.~1, pp.~67--78, 1980.

\bibitem{nelson1991} D.~B. Nelson, ``Conditional heteroskedasticity in asset returns: a new approach,'' \emph{Econometrica}, vol.~59, no.~2, pp.~347--370, 1991.

\bibitem{rogers1991} L.~C.~G. Rogers and S.~E. Satchell, ``Estimating variance from high, low and closing prices,'' \emph{Annals of Applied Probability}, vol.~1, no.~4, pp.~504--512, 1991.

\bibitem{glosten1993} L.~R. Glosten, R.~Jagannathan, and D.~E. Runkle, ``On the relation between the expected value and the volatility of the nominal excess return on stocks,'' \emph{Journal of Finance}, vol.~48, no.~5, pp.~1779--1801, 1993.

\bibitem{bns2004} O.~E. Barndorff-Nielsen and N.~Shephard, ``Power and bipower variation with stochastic volatility and jumps,'' \emph{Journal of Financial Econometrics}, vol.~2, no.~1, pp.~1--37, 2004.

\bibitem{corsi2009} F.~Corsi, ``A simple approximate long-memory model of realized volatility,'' \emph{Journal of Financial Econometrics}, vol.~7, no.~2, pp.~174--196, 2009.

\bibitem{hamilton1989} J.~D. Hamilton, ``A new approach to the economic analysis of nonstationary time series and the business cycle,'' \emph{Econometrica}, vol.~57, no.~2, pp.~357--384, 1989.

\bibitem{bucci2020} A.~Bucci, ``Realized volatility forecasting with neural networks,'' \emph{Journal of Financial Econometrics}, vol.~18, no.~3, pp.~502--531, 2020.

\bibitem{nystrup2021} P.~Nystrup, E.~Lindstr\"{o}m, and H.~Madsen, ``Hyperparameter optimization for portfolio selection with regime-switching models,'' \emph{Expert Systems with Applications}, vol.~184, p.~115538, 2021.

\bibitem{lim2021} B.~Lim, S.~\"{O}.~Ar\i k, N.~Loeff, and T.~Pfister, ``Temporal Fusion Transformers for interpretable multi-horizon time series forecasting,'' \emph{International Journal of Forecasting}, vol.~37, no.~4, pp.~1748--1764, 2021.

\bibitem{kim2022} H.~Y. Kim and C.~H. Won, ``Forecasting financial volatility with attention-based recurrent neural networks,'' \emph{Expert Systems with Applications}, vol.~198, p.~116879, 2022.

\bibitem{wu2023} J.~Wu, Y.~Chen, T.~Zhou, and L.~Li, ``Transformer-based models for multi-step realized volatility forecasting,'' \emph{International Journal of Forecasting}, vol.~39, no.~4, pp.~1768--1783, 2023.

\bibitem{zhang2023} Y.~Zhang, J.~Li, and X.~Zhu, ``GARCH-informed neural networks for volatility prediction,'' \emph{Quantitative Finance}, vol.~23, no.~7--8, pp.~1125--1142, 2023.

\bibitem{rahimikia2023} E.~Rahimikia and S.-H.~Poon, ``Machine learning for realized volatility forecasting: the role of news and sentiment,'' \emph{Journal of Empirical Finance}, vol.~74, p.~101424, 2023.

\bibitem{christensen2023} K.~Christensen, M.~Siggaard, and B.~Veliyev, ``A machine learning approach to volatility forecasting,'' \emph{Journal of Financial Econometrics}, vol.~21, no.~5, pp.~1680--1727, 2023.

\bibitem{olorunnimbe2023} M.~K. Olorunnimbe and H.~L. Viktor, ``Deep learning in stock market forecasting: a comprehensive survey,'' \emph{Neural Computing and Applications}, vol.~35, pp.~9947--9993, 2023.

\bibitem{chen2024} Z.~Chen, M.~Li, and H.~Wang, ``Temporal Fusion Transformers for cryptocurrency volatility,'' \emph{Finance Research Letters}, vol.~60, p.~104892, 2024.

\bibitem{patel2024} R.~Patel and A.~Sharma, ``Volatility forecasting in emerging markets using hybrid deep learning with regime awareness,'' \emph{Applied Soft Computing}, vol.~151, p.~111153, 2024.

\bibitem{zhao2024} W.~Zhao, Y.~Liu, and J.~Chen, ``Attention-based multi-scale volatility prediction with high-frequency data,'' \emph{IEEE Transactions on Neural Networks and Learning Systems}, vol.~35, no.~3, pp.~3891--3905, 2024.

\end{thebibliography}
}
