% Dvision.tex (safe include)
% Note: no \cleardoublepage here; let the main document control page breaks.

\clearpage
\thispagestyle{empty}

\begingroup
\centering

{\Large\bfseries Department Vision}\par\vspace{0.75em}
\emph{``To fortify Ethical Computational Excellence''}\par\vspace{1.5em}

{\Large\bfseries Department Mission}\par\vspace{0.75em}
\begin{minipage}{0.9\textwidth}
\raggedright
\begin{enumerate}
\item[M1:] Imparts core and contemporary knowledge in the areas of Computation and Information Technology.
\item[M2:] Promotes the culture of research and facilitates higher studies.
\item[M3:] Acquaints the students with the latest industrial practices, team building and entrepreneurship.
\item[M4:] Sensitizes the students to serve for environmental, social \& ethical needs of society through lifelong learning.
\end{enumerate}
\end{minipage}\par\vspace{1.5em}

{\Large\bfseries Program Educational Objectives (PEOs)}\par\vspace{0.75em}
\begin{minipage}{0.9\textwidth}
\raggedright
\begin{enumerate}
\item[PEO1:] \textbf{Professional Acumen:} Understand, analyze and design solutions with professional competency for real world problems.
\item[PEO2:] \textbf{Critical Analysis:} Develop software/embedded solutions for the requirements, based on critical analysis and research.
\item[PEO3:] \textbf{Team work:} Function effectively in a team and as an individual in a multidisciplinary/multicultural environment.
\item[PEO4:] \textbf{Life Long Learning:} Accomplish holistic development comprehending professional responsibilities.
\end{enumerate}
\end{minipage}

\endgroup
\clearpage
