% All the Variable should be entered inside {}
% If not applicable leave empty {} do not delete {}.

% Name of the University
\newcommand{\UniversityName}
{CHRIST (Deemed to be University)}


% Title of the project or thesis in capital letters
\newcommand{\ProjectTitle}
{REGIME-AWARE MULTI-ASSET VOLATILITY FORECASTING USING TEMPORAL FUSION TRANSFORMERS}


% Title of the project or thesis in title case
\newcommand{\ProjectTitleTwo}
{Regime-Aware Multi-Asset Volatility Forecasting using Temporal Fusion Transformers}


% Type of Dissertation M.Tech Dissertation or B.Tech Dissertation
\newcommand{\DissertationType}
{A Project Report on}

% Program name: MASTER OF TECHNOLOGY or BACHELOR OF TECHNOLOGY in Uppercase
\newcommand{\ProgramNameUpper}
{MASTER OF TECHNOLOGY}

% Program name: Master of Technology or Bachelor of Technology in Normal case
\newcommand{\ProgramName}
{Master of Technology}

% Program Short form: M.Tech. or B.Tech.
\newcommand{\ProgramNameShort}
{M.Tech.}


% Specialization
\newcommand{\SpecializationName}
{AI and Data Science Engineering}


% Name of Student
\newcommand{\AuthorNameOne}
{Abhishek Deep}

\newcommand{\RegisterNoOne}
{2567201}

\newcommand{\AuthorDepartmentOne}
{AI and Data Science Engineering}


\newcommand{\AuthorNameTwo}
{}

\newcommand{\RegisterNoTwo}
{}

\newcommand{\AuthorDepartmentTwo}
{}


\newcommand{\AuthorNameThree}
{}

\newcommand{\RegisterNoThree}
{}

\newcommand{\AuthorDepartmentThree}
{}



\newcommand{\AuthorDepartmentFive}
{}



% Name of the Guide
\newcommand{\GuideName}
{Dr. Alpha Vijayan}

% Designation of the Guide
\newcommand{\GuideDesignation}
{Assistant Professor}

% Department of Guide where he belongs to
\newcommand{\GuideDepartment}
{Department of AI and Data Science Engineering}


% Name of the Co-Guide
\newcommand{\CoGuideName}
{Dr. Michael Moses T}

% Designation of the Co-Guide
\newcommand{\CoGuideDesignation}
{Associate Professor}

% Department of Co-Guide where he belongs to
\newcommand{\CoGuideDepartment}
{Department of AI and Data Science Engineering}


\newcommand{\NameofAuthortobeCertified}
{Abhishek Deep}


% Address of the department where project officially executed
\newcommand{\DepartmentNameAddress}
{ %\renewcommand{\baselinestretch}{1.0}
\textbf{\large{
	\DepartmentName\\ 
	\CollegeName, \\ \UniversityName,\\ 
	Kumbalgodu,\,Bengaluru\,-\,560~074
}}

} 


% Project Date in month-yyyy format
\newcommand{\ProjectDate}
{March-2026}

% Academic Year in yyyy-yyyy format
\newcommand{\AcademicYear}
{2025-2026}

% Name of the department only
\newcommand{\DepartmentName}
{AI and Data Science Engineering}

% Name of the College
\newcommand{\CollegeName}
{School of Engineering and Technology}


% Name of the HOD or Coordinator
\newcommand{\HODorCoordinator}
{Dr Michael Moses T}

% Name of the Position of the Head
\newcommand{\HODorCoordinatorDesignation}
{Department Coordinator}


% Name of the Vice-Chancellor
\newcommand{\VCName}
{Dr Rev Fr Joseph C C}

% Name of the Pro Vice Chancellor
\newcommand{\ProVCName}
{Dr Rev Fr Viju P D}

% Name of the Director of the College
\newcommand{\DirectorName}
{Dr Fr Jiby Jose}

% Name of the Assistant Director of the College
\newcommand{\AssistantDirectorName}
{Fr Shijin P J}


% Name of the Dean/Associate Dean
\newcommand{\DeanName}
{Dr Raghunandan Kumar R}

% Name of the Position of Dean/Associate Dean
\newcommand{\DeanDesignation}
{Dean}



% Name of the Dean/Associate Dean
\newcommand{\AssociateDeanName}
{Dr E A Mary Anita }

% Name of the Position of Dean/Associate Dean
\newcommand{\AssociateDeanDesignation}
{Associate Dean}

% 'We' or 'I'
\newcommand{\WeorI}
{I}

% Acknowledgement for Guide
\newcommand{\AcknowledgementForGuide}
{
	\WeorI~am deeply grateful to my guide, \textbf{\GuideName}, for her constant support, encouragement, and expert guidance throughout the course of this project. Her valuable insights in machine learning and data science shaped every aspect of this work.
}

% Acknowledgement for Co-Guide (Leave blank if no co-guide)
\newcommand{\AcknowledgementForCoGuide}
{}


% Acknowledgement for Others
\newcommand{\AcknowledgementForOthers}
{
	\WeorI~would also like to express heartfelt gratitude to my parents for their unwavering support, encouragement, and blessings throughout this journey. \WeorI~thank all the faculty members and non-teaching staff of the department for their help and cooperation.
}

% Date of Declaration
\newcommand{\DateofDeclaration}
{14-03-2026}

% Department Vision
\newcommand{\DepartmentVision}
{To excel in Human-Centred AI and Data-Driven Innovation
}


% Department Mission
\newcommand{\DepartmentMission}
{
			\textbf{M1: }  Empowering individuals to ethically harness data and AI through accessible and value-driven curriculum. \\
			\textbf{M2: }  Foster a dynamic research environment that advances innovative and impactful solutions for the betterment of global well-being. \\
			\textbf{M3: }  Innovate scientific knowledge and entrepreneurship through academia and Industry collaborations.  \\

}


% Department PEOs
\newcommand{\DepartmentPEOs}
{
	\textbf{PEO1:} Professional Acumen : Understand, analyze and design solutions with professional competency for the real world problems. \\
	\textbf{PEO2:} Critical Analysis : Develop software/embedded solutions for the requirements, based on critical analysis and research.\\
	\textbf{PEO3:} Team work : Function effectively in a team and as an individual in a multidisciplinary/ multicultural environment.\\
	\textbf{PEO4:} Life Long Learning : Accomplish holistic development comprehending professional responsibilities\\
}

% Department PSOs
\newcommand{\DepartmentPSOs}
{
	\textbf{Artificial Intelligence:} To utilize artificial intelligence principles to solve real world problems \\
	\textbf{Machine Learning:} To select suitable machine learning algorithms for the given problem\\
	\textbf{Service Learning:} Analyze Social Relevant Problems and design solutions through Service Learning\\
}
